---
title: "A Shiny Starter"
author: "Andrew Duncan"
date: ""
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, comment = NA, message = F)
library(dplyr)
library(knitr)
```

## What is Shiny?

```{r app_1st_time, echo = FALSE}
shinyApp(ui = shinyUI(fluidPage(
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30)
      ),
      # plot and text output
      mainPanel(
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
)),

# Define server logic required to draw a histogram
server = function(input, output) {
   output$disPlot = renderPlot({
      # generate bins based on input$bins from ui.R
      x    = faithful[, 2] 
      bins = seq(min(x), max(x), length.out = input$bins + 1)
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
   output$disText = renderText({
     paste("Number of bins:", input$bins)
   })
})
```

## What is Shiny?

>- `shiny` package
>- Turns `R` into _html_, _css_ and _javascript_.
>- Can customise and style it. 
>- If you can do it in `R`, can do it in shiny.
>- Might not be a good idea though.
>- Best apps I've seen are plot/image based.

## Should I be interested?

- Makes `R` interactive (esp. plotting and data)
- Can show your work online
    + shiny server
    + shinyapps.io (free and paid)
- Excellent way to share work.

## How does it work?

- Either two files (`ui.R` and `server.R`)
- or one file `app.R` which contains both `ui` and `server`.
- `ui` the _user interface_
    + how the app will look (style and locations)
    + the inputs (sliders, textboxes, dropdowns)
    + the outputs (plots, tables, text)
- `server` the server side functions
    + what the user doesn't see (`R` code)
    + takes the _inputs_ and produces the _outputs_
    
## An Example

```{r app_ui, eval = F}
ui = shinyUI(fluidPage(
  # App title
  titlePanel("Old Faithful Geyser Data"),
  # this app has sidebar for inputs
  sidebarLayout(
      sidebarPanel(
        # one input - a slider
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30)
         ),
  # with a main panel for the outputs    
      mainPanel(
        # two outputs - a plot and some text
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
))
```

## An Example

```{r app_server, eval = F}
server = shinyServer(function(input, output) {
  # this output is a plot
   output$disPlot = renderPlot({
      x    = faithful[, 2] 
      # we access the inputs via input$inputId
      bins = seq(min(x), max(x), length.out = input$bins + 1)
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
   # the second output is some text
   output$disText = renderText({
     paste("Number of bins:", input$bins)
   })
})
```

## The App

```{r app_2nd_time, echo = FALSE}
shinyApp(ui = shinyUI(fluidPage(
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30)
      ),
      # plot and text output
      mainPanel(
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
)),

# Define server logic required to draw a histogram
server = function(input, output) {
   output$disPlot = renderPlot({
      # generate bins based on input$bins from ui.R
      x    = faithful[, 2] 
      bins = seq(min(x), max(x), length.out = input$bins + 1)
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
   output$disText = renderText({
     paste("Number of bins:", input$bins)
   })
})
```


## Inputs, Outputs and Layouts

- There are a wide range of inputs and outputs. 
- Different styles are built in or can build from scratch.

```{r shiny_io, echo = FALSE, results = 'asis'}
shiny_io = data_frame(
  inputs = c("Check box, Radio Button","Dates, Numeric  or Text",
             "File Upload", "Slider", "Drop down (select)"),
  outputs = c("HTML", "Plots", "Tables", "Files to download", "Text"))
kable(shiny_io)
```


## Some Examples

```{r navbar, echo = F}
shinyApp(ui = shinyUI(navbarPage("Navbar!",
  tabPanel("Plot",
    sidebarLayout(
      sidebarPanel(
        radioButtons("plotType", "Plot type",
          c("Scatter"="p", "Line"="l")
        )
      ),
      mainPanel(
        plotOutput("plot")
      )
    )
  ),
  tabPanel("Summary",
    verbatimTextOutput("summary")
  ),
  navbarMenu("More",
    tabPanel("Table",
      DT::dataTableOutput("table")
    )
  )
)),

server = shinyServer(function(input, output, session) {
  output$plot <- renderPlot({
    plot(cars, type=input$plotType)
  })

  output$summary <- renderPrint({
    summary(cars)
  })

  output$table <- DT::renderDataTable({
    DT::datatable(cars)
  })
}))
```



## Some Examples

```{r tabs, echo = F}
shinyApp(ui = # Define UI for random distribution application 
shinyUI(fluidPage(
    
  # Application title
  titlePanel("Tabsets"),
  
  # Sidebar with controls to select the random distribution type
  # and number of observations to generate. Note the use of the
  # br() element to introduce extra vertical spacing
  sidebarLayout(
    sidebarPanel(
      radioButtons("dist", "Distribution type:",
                   c("Normal" = "norm",
                     "Uniform" = "unif",
                     "Log-normal" = "lnorm",
                     "Exponential" = "exp")),
      br(),
      
      sliderInput("n", 
                  "Number of observations:", 
                   value = 500,
                   min = 1, 
                   max = 1000)
    ),
    
    # Show a tabset that includes a plot, summary, and table view
    # of the generated distribution
    mainPanel(
      tabsetPanel(type = "tabs", 
        tabPanel("Plot", plotOutput("plot")), 
        tabPanel("Summary", verbatimTextOutput("summary")), 
        tabPanel("Table", tableOutput("table"))
      )
    )
  )
)),
server = # Define server logic for random distribution application
shinyServer(function(input, output) {
  
  # Reactive expression to generate the requested distribution.
  # This is called whenever the inputs change. The output
  # functions defined below then all use the value computed from
  # this expression
  data <- reactive({
    dist <- switch(input$dist,
                   norm = rnorm,
                   unif = runif,
                   lnorm = rlnorm,
                   exp = rexp,
                   rnorm)
    
    dist(input$n)
  })
  
  # Generate a plot of the data. Also uses the inputs to build
  # the plot label. Note that the dependencies on both the inputs
  # and the data reactive expression are both tracked, and
  # all expressions are called in the sequence implied by the
  # dependency graph
  output$plot <- renderPlot({
    dist <- input$dist
    n <- input$n
    
    hist(data(), 
         main=paste('r', dist, '(', n, ')', sep=''))
  })
  
  # Generate a summary of the data
  output$summary <- renderPrint({
    summary(data())
  })
  
  # Generate an HTML table view of the data
  output$table <- renderTable({
    data.frame(x=data())
  })
  
})
)
```


## Some Examples

```{r conditional, echo = F}
shinyApp(ui = shinyUI(fluidPage(

  titlePanel("Conditional panels"),

  column(4, wellPanel(
    sliderInput("n", "Number of points:",
                min = 10, max = 200, value = 50, step = 10)
  )),

  column(5,
    "The plot below will be not displayed when the slider value",
    "is less than 50.",

    # With the conditionalPanel, the condition is a JavaScript
    # expression. In these expressions, input values like
    # input$n are accessed with dots, as in input.n
    conditionalPanel("input.n >= 50",
      plotOutput("scatterPlot", height = 300)
    )
  )
)),

server = shinyServer(function(input, output) {

  output$scatterPlot <- renderPlot({
    x <- rnorm(input$n)
    y <- rnorm(input$n)
    plot(x, y)
  })

})
)
```

## Some Examples


```{r dynamic_ui, echo = F}
shinyApp(ui = shinyUI(fluidPage(
  titlePanel("Dynamically generated user interface components"),
  fluidRow(

    column(4, wellPanel(
      selectInput("input_type", "Input type",
        c("slider", "text", "numeric", "checkbox",
          "checkboxGroup", "radioButtons", "selectInput",
          "selectInput (multi)", "date", "daterange"
        )
      )
    )),

    column(4, wellPanel(
      # This outputs the dynamic UI component
      uiOutput("ui")
    )),

    column(4,
      tags$p("Input type:"),
      verbatimTextOutput("input_type_text"),
      tags$p("Dynamic input value:"),
      verbatimTextOutput("dynamic_value")
    )
  )
)),

server = shinyServer(function(input, output) {

  output$ui <- renderUI({
    if (is.null(input$input_type))
      return()

    # Depending on input$input_type, we'll generate a different
    # UI component and send it to the client.
    switch(input$input_type,
      "slider" = sliderInput("dynamic", "Dynamic",
                             min = 1, max = 20, value = 10),
      "text" = textInput("dynamic", "Dynamic",
                         value = "starting value"),
      "numeric" =  numericInput("dynamic", "Dynamic",
                                value = 12),
      "checkbox" = checkboxInput("dynamic", "Dynamic",
                                 value = TRUE),
      "checkboxGroup" = checkboxGroupInput("dynamic", "Dynamic",
        choices = c("Option 1" = "option1",
                    "Option 2" = "option2"),
        selected = "option2"
      ),
      "radioButtons" = radioButtons("dynamic", "Dynamic",
        choices = c("Option 1" = "option1",
                    "Option 2" = "option2"),
        selected = "option2"
      ),
      "selectInput" = selectInput("dynamic", "Dynamic",
        choices = c("Option 1" = "option1",
                    "Option 2" = "option2"),
        selected = "option2"
      ),
      "selectInput (multi)" = selectInput("dynamic", "Dynamic",
        choices = c("Option 1" = "option1",
                    "Option 2" = "option2"),
        selected = c("option1", "option2"),
        multiple = TRUE
      ),
      "date" = dateInput("dynamic", "Dynamic"),
      "daterange" = dateRangeInput("dynamic", "Dynamic")
    )
  })

  output$input_type_text <- renderText({
    input$input_type
  })

  output$dynamic_value <- renderPrint({
    str(input$dynamic)
  })

}))
```

## Reactive Programming

- `shiny` is built upon _reactive expressions_.
- inputs and outputs are reactive. 
- _reactive expressions_ aren't like regular `R`. They listen!
- Built in or create your own. 


## Reactive Programming

- Problem is when a function takes a long time.
- Need to control the listening.
- Can use `isolate()` with `actionButton()` or `submitButton`.


## `isolate` with `actionButton()`

- Can choose to `isolate` one or more inputs, then listen for `actionButton()`.
- Add an `actionButton()` to the `ui`.

```{r isolate_ui, eval = F}
ui = shinyUI(fluidPage(
   titlePanel("Old Faithful Geyser Data"),
   sidebarLayout(
      sidebarPanel(
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30),
         # add an action button
         actionButton(inputId = "plot", label = "Action Button")
         ),
      mainPanel(
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
))
```

## `isolate` with `actionButton()`

- In `server` we `isolate` the inputs we want to but listen to the `actionButton`.

```{r isolate_server, eval = F}
server = shinyServer(function(input, output) {
   output$disPlot = renderPlot({
     input$plot # listen to the actionButton
      x    = faithful[, 2] 
      isolate({ # isolate the plot which depend on the 'bins' input
        bins = seq(min(x), max(x), length.out = input$bins + 1)
        hist(x, breaks = bins, col = 'darkgray', border = 'white')
      })
   })
   output$disText = renderText({
     # chosen not to isolate the text, it will still listen.
     paste("Number of bins:", input$bins)
   })
})
```


## The App

```{r app_3rd_time, echo = FALSE}
shinyApp(ui = shinyUI(fluidPage(
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30),
         # add an action button
         actionButton(inputId = "plot", label = "Action Button")
      ),
      # plot and text output
      mainPanel(
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
)),

# Define server logic required to draw a histogram
server = function(input, output) {
   output$disPlot = renderPlot({
     input$plot # listen to the actionButton
      x    = faithful[, 2] 
      isolate({ # isolate the plot which depend on the 'bins' input
        bins = seq(min(x), max(x), length.out = input$bins + 1)
        hist(x, breaks = bins, col = 'darkgray', border = 'white')
      })
   })
   output$disText = renderText({
     # chosen not to isolate the text, it will still listen.
     paste("Number of bins:", input$bins)
   })
})
```

## The `submitButton()`

- The alternative is to add a `submitButton()`.
- This causes everything to wait. 
- Just change the `ui`. 

```{r submit_ui, eval = F}
ui = shinyUI(fluidPage(
   titlePanel("Old Faithful Geyser Data"),
   sidebarLayout(
      sidebarPanel(
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30),
         # add a submit button
         submitButton(text = "Submit Button")
         ),
      mainPanel(
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
))
```

## The App
```{r app_4th_time, echo = FALSE}
shinyApp(ui = shinyUI(fluidPage(
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput(inputId = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30),
         submitButton(text = "Submit Button")
      ),
      # plot and text output
      mainPanel(
         plotOutput(outputId = "disPlot"),
         textOutput(outputId = "disText")
      )
   )
)),

# Define server logic required to draw a histogram
server = function(input, output) {
   output$disPlot = renderPlot({
      # generate bins based on input$bins from ui.R
      x    = faithful[, 2] 
      bins = seq(min(x), max(x), length.out = input$bins + 1)
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
   output$disText = renderText({
     paste("Number of bins:", input$bins)
   })
})
```


## Summary

- `shiny` takes `R` and makes it online and interactive.
- (imo) Best suited to plots or tabulated data investigation.
- Built on _reactivity_, expressions that listen!
- _reactivity_ has problems with slow functions.
- Can slow/stop the listening if you want.
- [shiny.rstudio.com](http://shiny.rstudio.com)

## Thanks for Listening

### Any Questions?

