---
title: "A Starter for Shiny"
author: "Andrew Duncan"
date: ""
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, comment = NA, message = F)
```

## What is Shiny?

```{r, echo = FALSE}
shinyApp(ui = shinyUI(fluidPage(
   
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput("bins",
                     "Number of bins:",
                     min = 1,
                     max = 50,
                     value = 30)
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("disPlot"),
         textOutput("disText")
      )
   )
)),

# Define server logic required to draw a histogram
server = function(input, output) {
   
   output$disPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2] 
      bins <- seq(min(x), max(x), length.out = input$bins + 1)
      
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
      
      
   })
   output$disText <- renderText({
     paste("Number of bins:", input$bins)
   })
})
```

## What is Shiny?


## Should I be interested?

- Sharing
- Interactive
- Online


## How does it work?

- `shiny` package
- Turns `R` into _html_, _css_ and _javascript_.
- Can customise and style it. 

## Sharing

- Online
    + shiny server
    + shinyapps.io
- Share the code directly

## Some technical details

- Two parts. `ui` and `server`.
- `ui` contains the things the user sees and can change.
- Split into _inputs_ and _outputs_.

```{r, eval = F}
ui <- shinyUI(fluidPage(
   titlePanel("Old Faithful Geyser Data"),
   sidebarLayout(
      sidebarPanel(
         sliderInput(id = "bins", label = "Number of bins:",
                     min = 1, max = 50, value = 30)
         ),
      mainPanel(
         plotOutput("disPlot"),
         textOutput("disText")
      )
   )
))
```

## Some technical details

- `server` contains the things the user doesn't see (your `R` code).
- needs function dependent on _input_ and _output_.
- specific functions to _render_ the output that the user sees.

```{r, eval = F}
server <- shinyServer(function(input, output) {
   output$disPlot <- renderPlot({
      x    <- faithful[, 2] 
      bins <- seq(min(x), max(x), length.out = input$bins + 1)
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
   output$disText <- renderText({
     paste("Number of bins:", input$bins)
   })
})
```

## The App

```{r, echo = FALSE}
shinyApp(ui = shinyUI(fluidPage(
   
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput("bins",
                     "Number of bins:",
                     min = 1,
                     max = 50,
                     value = 30)
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("disPlot"),
         textOutput("disText")
      )
   )
)),

# Define server logic required to draw a histogram
server = function(input, output) {
   
   output$disPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2] 
      bins <- seq(min(x), max(x), length.out = input$bins + 1)
      
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
      
      
   })
   output$disText <- renderText({
     paste("Number of bins:", input$bins)
   })
})
```

